Erosion of somatic tissue identity with loss of the X-linked intellectual disability factor KDM5C

### 2024-07-02 submission/manuscript.Rmd
remmoved commas from GO IDs by typing by hand like a peasant

	if (cluster == "no") {
		out <- df_test[df_test$Description == x,][column][,1]
	} else {
		out <- df_test[df_test$Description == x & df_test$Cluster == cluster,][column][,1]
	}
	return(out)
}

@@

The testis contains both germ cells (meiotic cells, e.g. spermatogonia) and somatic cells (non-meiotic, e.g. Leydig cells) that support hormone production and germline functions. Select testis-enriched DEGs that were previously characterized had germline-specific functions[], suggesting _Kdm5c_-KO cells fail to demarcate between the soma and germline. To test if this holds true for all _Kdm5c_-KO testis-biased DEGs, we first assed their function through gene onotology. We found high enrichment of germline-relevant ontologies, including spermatid development (`r getGO("spermatid development", testisGO, 'ID', "no")`, p.adjust = `r signif(getGO("spermatid development", testisGO, 'p.adjust', "no"), 3)`) and sperm axoneme assembly (`r getGO("sperm axoneme assembly", testisGO, 'ID', "no")`, p.adjust = `r signif(getGO("sperm axoneme assembly", testisGO, 'p.adjust', "no"), 3)`) (Figure 2A).

@@

To evaluate if all germline DEGs are constitutively dysregulated or change over the course of development, we then compared EpiLC germline DEGs to those expressed in the _Kdm5c_-KO brain. The majority of EpiLC germline DEGs were unique to EpiLCs, with only _Cyct_ shared across all sequencing datasets (Figure 3F). We then compared the function of EpiLC and brain germline DEGs and found particularly high enrichment of meiosis-related gene ontologies in EpiLCs (Figure 3G), such as meiotic cell cycle (`r getGO("meiotic cell cycle", GO_EpiBrain, 'ID', "EpiLC")`, p.adjust = `r signif(getGO("meiotic cell cycle", GO_EpiBrain, 'p.adjust', "EpiLC"), 3)`). While there was modest enrichment of meiotic gene ontologies in in both brain regions, the _Kdm5c_-KO hippocampus showed strong enrichment of late-stage sperm genes, such those involved in the sperm axoneme. This shift from meiotic genes to later spermatogenesis genes in the hippocampus suggests the germline developmental program could occur ectopically as _Kdm5c_-KO cells progresses through somatic tissue development. **note: this is strengthened by the ChIP-seq data since KDM5C is not directly bound to many brain/flagellar DEGs.This point might be stronger in the ChIPseq figure**

>>>>

	if (cluster == "no") {
		out <- df_test[df_test$Description == x,][column][,1]
	} else {
		out <- df_test[df_test$Description == x & df_test$Cluster == cluster,][column][,1]
	}
	outformat <- ifelse(column == "ID", toString(out), out) #an attempt to remove the commas from GO ID numbers, didn't work
	return(outformat) 
}

@@

The testis contains both germ cells (meiotic cells, e.g. spermatogonia) and somatic cells (non-meiotic, e.g. Leydig cells) that support hormone production and germline functions. Select testis-enriched DEGs that were characterized previously had germline-specific functions[@scandagliaLossKdm5cCauses2017], suggesting _Kdm5c_-KO cells fail to demarcate between the soma and germline. To test if this holds true for all _Kdm5c_-KO testis-biased DEGs, we first assed their function through gene onotology. We found high enrichment of germline-relevant ontologies, including spermatid development (GO: 0007286, p.adjust = `r signif(getGO("spermatid development", testisGO, 'p.adjust', "no"), 3)`) and sperm axoneme assembly (GO: 0007288, p.adjust = `r signif(getGO("sperm axoneme assembly", testisGO, 'p.adjust', "no"), 3)`) (Figure 2A).

@@

To evaluate if all germline DEGs are constitutively dysregulated or change over the course of development, we then compared EpiLC germline DEGs to those expressed in the _Kdm5c_-KO brain. The majority of EpiLC germline DEGs were unique to EpiLCs, with only _Cyct_ shared across all sequencing datasets (Figure 3F). We then compared the function of EpiLC and brain germline DEGs and found particularly high enrichment of meiosis-related gene ontologies in EpiLCs (Figure 3G), such as meiotic cell cycle (GO: 0051321, p.adjust = `r signif(getGO("meiotic cell cycle", GO_EpiBrain, 'p.adjust', "EpiLC"), 3)`). While there was modest enrichment of meiotic gene ontologies in in both brain regions, the _Kdm5c_-KO hippocampus showed strong enrichment of late-stage sperm genes, such those involved in the sperm axoneme. This shift from meiotic genes to later spermatogenesis genes in the hippocampus suggests the germline developmental program could occur ectopically as _Kdm5c_-KO cells progresses through somatic tissue development. **note: this is strengthened by the ChIP-seq data since KDM5C is not directly bound to many brain/flagellar DEGs.This point might be stronger in the ChIPseq figure**
